---
title: "Survival Data Analysis HW #1"
author: "Brendan Graham"
date: "June 20, 2016"
output: html_document
---

```{r setup, include=FALSE}
library(survival)
library(OIsurv)
library(KMsurv)
library(km.ci)
library(locfit)
library(dplyr)
library(survMisc)

#load data 
wha <- read.table(file = "https://raw.githubusercontent.com/brndngrhm/survival_analysis/master/data/whas500.txt", 
                  header = T)
```

### 2.
#### a)
 
```{r, echo=T, fig.height=8, fig.width=8}
wha.surv <- Surv(wha$lenfol, wha$fstat)
fit <- survfit(wha.surv~1)
surv.plot <- plot(fit, conf.int = T, xlab = "Number of days followed since MI",
             ylab = "Survival Probability", main = "K-M Survival Curve of Heart Attack Survivors")
conf.bands <- km.ci(fit, method="loghall")
lines(conf.bands, lty=3)
legend(500, .4, c("K-M Curve", "Pointwise CI", "Simultaneous CB"), lty=c(1,2,3))
```

#### b) 
The estimated survival curve does not appears to be headed towards zero as time goes to infinity. The curve seems to be levelling-off around .40 to .50 at around 1,500 days.


#### c)

```{r, echo=T}
quantile(fit, probs = c(.25, .5, .75))
```

To obtain the 95% confidence interval, R looks for every value of t such that the point-wise confidence interval for S(t) includes 0.50. In this case, the lower bound is t=1527, which has a 95% CI of (0.4958, 0.598). The next lowest time, t=1506 has a 95% CI of (0.5020, 0.602) which does not include 0.50 which is why the lower bound stopped at t=1527. Every CI for values of t > 1527 includes 0.50 which is why there is no upper bound reported by R.

#### d)
```{r, echo=T}
#split into men and women dataframes
men <- wha %>% filter(gender == 0)
women <- wha %>% filter(gender == 1)

#mean survival time for men
men.surv <- Surv(subset(men$lenfol,men$gender == 0), subset(men$fstat, men$gender == 0))
fit.men <- survfit(men.surv~1)
print(fit.men, print.rmean =T)

#mean survival time for women
women.surv <- Surv(subset(women$lenfol,women$gender == 1), subset(women$fstat, women$gender == 1))
fit.women <- survfit(women.surv~1)
print(fit.women, print.rmean =T)

#mrlife function
mrlife <- function(t, event, censoring){
  kmfit <- survfit(Surv(event[event>t]-t, censoring[event>t])~1, conf.type="log-log")
  print(kmfit, print.rmean = T)
}

#Mean Residual lifetime for Men at time 1000
mrlife(1000,men$lenfol, men$fstat)

#Mean Residual lifetime for Women at time 1000 
mrlife(1000,women$lenfol, women$fstat)



```

#### e) 
```{r, echo=T, fig.height=8, fig.width=8}
plot(fit.men, conf.int = F, main="Survival Curves for Men and Women", ylab="Survival probability", xlab="Number of days followed since MI")
lines(fit.women, conf.int = F, lwd=2)
legend(500, .4, c("Male", "Female"), lwd = c(1,2))
```

#### f)
Ho: S~male~(t) = S~female~(t)  
Ha: S~male~(t) != S~female~(t)
```{r, echo=T}
fit <- ten(wha.surv ~ gender, data=wha)
comp(fit)
```
Since the survival curve for men and women do not cross we can just look at the results in the first table. Results in the first table from the 6 tests all show a signfificant difference between at at least the .05 level. Therefore we reject the null and conclude the alternative. We have enough evidence to conclude there is a difference between the survial curves for men and women. Based ont he survival plots in part (e), males have the higher survival probability.

#### g)
