---
title: 'Survival Data Analysis HW #1'
author: "Brendan Graham"
date: "June 20, 2016"
output: html_document
---

```{r setup, include=FALSE}
library(survival)
library(OIsurv)
library(KMsurv)
library(km.ci)
library(locfit)
library(dplyr)
library(survMisc)

#load data 
wha <- read.table(file = "https://raw.githubusercontent.com/brndngrhm/survival_analysis/master/data/whas500.txt", 
                  header = T)
wha <- wha %>% select(age, gender, lenfol, fstat)
```

### 2.
#### a)
 
```{r, echo=T, fig.height=8, fig.width=8}
wha.surv <- Surv(wha$lenfol, wha$fstat)
fit <- survfit(wha.surv~1, conf.type="log-log")
summary(fit)
surv.plot <- plot(fit, conf.int = T, xlab = "Number of days followed since MI",
             ylab = "Survival Probability", main = "K-M Survival Curve of Heart Attack Survivors")
conf.bands <- km.ci(fit, method="loghall")
lines(conf.bands, lty=3)
legend(500, .4, c("K-M Curve", "Pointwise CI", "Simultaneous CB"), lty=c(1,2,3))
```

#### b) 
The estimated survival curve does head towards zero as time goes to infinity. Since every subject msu eventually die, the survival curve will head to zero, although at a slow rate.


#### c)
```{r, echo=T}
quantile(fit, probs = c(.25, .5, .75))
```

To obtain the 95% confidence interval for the 50th percentile, R looks for every value of t such that the point-wise confidence interval for S(t) includes 0.50. In this case, the lower bound is t=1527, which has a 95% CI of (0.4958, 0.598). The next lowest time, t=1506 has a 95% CI of (0.5020, 0.602), which does not include 0.50, so the lower bound stopped at t=1527. Every CI for values of t > 1527 includes 0.50 which is why there is no upper bound reported by R for the 95% confidence interval for the 50th percentile.

#### d)
The mean survival time for men is 1448.5 and the mean survival time for women is 1260.2.  

```{r, echo=T}
#split into men and women dataframes
men <- wha %>% filter(gender == 0)
women <- wha %>% filter(gender == 1)

#mean survival time for men
men.surv <- Surv(men$lenfol, men$fstat)
fit.men <- survfit(men.surv~1)
print(fit.men, print.rmean =T)

#mean survival time for women
women.surv <- Surv(women$lenfol, women$fstat)
fit.women <- survfit(women.surv~1)
print(fit.women, print.rmean =T)
```

At t=1000, the mean residual lifetime for men is  1060.4 and for women, it is 1063.8.
```{r, echo=T}
#mrlife function
mrlife <- function(t, event, censoring){
  kmfit <- survfit(Surv(event[event>t]-t, censoring[event>t])~1, conf.type="log-log")
  print(kmfit, print.rmean = T)
}

#Mean Residual lifetime for Men at time 1000
mrlife(1000, wha$lenfol[wha$gender == 0], wha$fstat[wha$gender == 0])

#Mean Residual lifetime for Women at time 1000 
mrlife(1000, wha$lenfol[wha$gender == 1], wha$fstat[wha$gender == 1])
```

#### e)
The survival curves for men and women shows the survival curve for men is higher than he survival curve for women.
```{r, echo=T, fig.height=8, fig.width=8}
plot(fit.men, conf.int = F, main="Survival Curves for Men and Women", ylab="Survival probability", xlab="Number of days followed since MI", col="Blue", lwd=2)
lines(fit.women, conf.int = F, lwd=2, col="red")
legend(500, .4, c("Male", "Female"), lwd = c(1,2), col=c("Blue", "red"))
```

#### f)
Ho: S~male~(t) = S~female~(t)  
Ha: S~male~(t) != S~female~(t)
```{r, echo=T}
fit.gender <- survfit(wha.surv ~ gender, data=wha)
fit <- ten(fit.gender)
comp(fit)
```
Since the survival curve for standard and test treatments cross,  we can look at the results in the second table of the output. Results in the second table  for the log rank test show a significant difference between the male and females curves at at the .05 level. Therefore we reject the null and conclude the alternative. We have enough evidence to conclude there is a difference between the survival curves for men and women at the .05 level. Based on the survival plots in part (e), males have higher survival probability.

#### g)
Ho: S~50 or less~(t) = S~Between 50 and 75~(t) = S~75 or Greater~(t)  
Ha: At least 2 equalities in Ho are not true

```{r, echo=T, fig.height=8, fig.width=8}
wha$age2 <- 2
wha$age2[wha$age <= 50] <- 1
wha$age2[wha$age >= 75] <- 3

fit.age <- ten(survfit(Surv(wha$lenfol, wha$fstat) ~ age2, data=wha))
comp(fit.age, data = wha)
```
The results in the first table from the 6 tests all indicate very high chi-square values. Therefore we reject the null and conclude the alternative. We have enough evidence at the .05 level to conclude there is a difference between the survival curves for the three age groups (50 or less, Between 50 and 75, 75 or Greater). Since there is a difference in the curves we can conduct a trend test. (I tried to do this in R with the following code, but it did not work)

Ho: There is a trend present in the survival curves
Ha: There is not a trend present in the survival curves

```{r, echo=T}
wha2 <- wha %>% arrange(lenfol)
cbind(fit.age$cg, unique(cbind(wha2$lenfol,wha2$age2)))[1:10,]
comp(fit.age, scores=c(3, 2, 1), data=wha)
```

I also tried to use SAS< but was having problems with getting citric reciever to work on my home computer. Here's the SAS code I would use if I could get SAS to open:

proc lifetest data = wha2;
  lenfol lenfol*fstat(1);
  strata age2 /trend test = (logrank wilcoxon tarone peto modpeto);
run;


To visualize the trend, below is a plot of the three survival curves by age category. I tried to do this in R with the following v

```{r, echo=T, fig.height=8, fig.width=8}
wha$age2 <- "Between 50 and 75"
wha$age2[wha$age <= 50] <- "50 or less"
wha$age2[wha$age >= 75] <- "75 or Greater"

fit.low <- survfit(Surv(wha$lenfol[wha$age2 == "50 or less"], wha$fstat[wha$age2 == "50 or less"])~1, conf.type = "log-log")
fit.med <- survfit(Surv(wha$lenfol[wha$age2 == "Between 50 and 75"], wha$fstat[wha$age2 == "Between 50 and 75"])~1, conf.type = "log-log")
fit.high <- survfit(Surv(wha$lenfol[wha$age2 == "75 or Greater"], wha$fstat[wha$age2 == "75 or Greater"])~1, conf.type = "log-log")
plot(fit.high, conf.int = F, main="Heart Attack Survival Curves by Age Category", xlab="Number of days followed since MI", ylab="Survival Probability", lwd=2, col="purple")
lines(fit.med, conf.int = F, col="Red", lwd=2)
lines(fit.low, conf.int = F, col="Blue", lwd=2)
legend(500, .2, c("Greater than 75", "Between 50 and 75","50 or less"), lwd=c(2,2,2), col=c("Purple", "Red", "Blue"))
```

