---
title: "Survival Data Analysis HW #1"
author: "Brendan Graham"
date: "June 20, 2016"
output: html_document
---

```{r setup, include=FALSE}
library(survival)
library(OIsurv)
library(KMsurv)
library(km.ci)

#load data 
wha <- read.table(file = "https://raw.githubusercontent.com/brndngrhm/survival_analysis/master/data/whas500.txt", 
                  header = T)
```

### 2.
#### a)
 
```{r, echo=T, fig.height=8, fig.width=8}
wha.surv <- Surv(wha$lenfol, wha$fstat)
fit <- survfit(wha.surv~1)
surv.plot <- plot(fit, conf.int = T, xlab = "Number of days followed since MI",
             ylab = "Survival Probability", main = "K-M Survival Curve of Heart Attack Survivors")
conf.bands <- km.ci(fit, method="loghall")
lines(conf.bands, lty=3)
legend(500, .4, c("K-M Curve", "Pointwise CI", "Simultaneous CB"), lty=c(1,2,3))
```

#### b) 
The estimated survival curve does not appears to be headed towards zero as time goes to infinity. The curve seems to be levelling-off around .40 to .50 at around 1,500 days.


#### c)

```{r, echo=T}
quantile(fit, probs = c(.25, .5, .75))
```

To obtain the 95% confidence interval, R looks for every value of t such that the point-wise confidence interval for S(t) includes 0.50. In this case, the lower bound is t=1527, which has a 95% CI of (0.4958, 0.598). The next lowest time, t=1506 has a 95% CI of (0.5020, 0.602) which does not include 0.50 which is why the lower bound stopped at t=1527. Every CI for values of t > 1527 includes 0.50 which is why there is no upper bound reported by R.

#### d)

#### e)

#### f)

#### g)
