---
title: "HW 2"
author: "Brendan Graham"
date: "June 24, 2016"
output: word_document
---

```{r setup, include=FALSE}
library(survival)
library(OIsurv)
library(KMsurv)
library(km.ci)
library(locfit)
library(dplyr)
library(survMisc)

```

### 2.
#### a)
```{r, echo=T}
#mean using integration
mean.wei <- function(x,t,shape, scale) x*dweibull(x, shape, scale)
integrate(mean.wei, 0, Inf, shape=1.5, scale=100)

#mean resid lifetime at t=100
mean.resid.wei <- function(x,t,shape,scale) {(x-t)*dweibull(x,shape,scale)/(1-pweibull(t, shape, scale))}
integrate(mean.resid.wei, 100, Inf, t=100, shape=1.5, scale=100)
```

#### b) 
The 35th percentile for survival time using a Weibull distribution with shape=1.5 and scale=100 is:
```{r, echo=T}
as.numeric(format(qweibull(.35, 1.5, 100), digits = 4))
```

#### c)
Appoximate mean survival time using a random sample of 100,000 from a Weibull distribution with shape 1.5 and scale 100:
```{r, echo=T}
set.seed(624)
samp <- rweibull(100000, 1.5, 100)
mean(samp)
```

Appoximate mean residual life using a random sample of 100,000 from a Weibull distribution with shape 1.5 and scale 100:
```{r, echo=T}
mean(samp[samp>100]) - 100
```

Appoximate 35th percentile using a random sample of 100,000 from a Weibull distribution with shape 1.5 and scale 100:
```{r, echo=T}
quantile(samp, probs=c(.35))
```

The approximate mean, residual life at 100 and 35th percentile from the random sample are roughly the same as the corresponding values calculated in part (a).